---
title: 'Reproducible Research: Peer Assessment 1'
author: "Qinqing Liu"
date: "10/1/2016"
output: pdf_document
---

## Loading and preprocessing the data
```{r}
unzip(zipfile="activity.zip")
data <- read.csv("activity.csv", header = TRUE, sep = ',',
                 colClasses = c("numeric", "character",
                                "integer"))
```

## What is mean total number of steps taken per day?
```{r}
library(ggplot2)
library(dplyr)
library(lubridate)
data$date <- ymd(data$date)

steps <- data %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps))
  
ggplot(steps, aes(x = steps)) +
  geom_histogram(binwidth = 1000) +
  labs(title = "Histogram of Steps per day", x = "Steps per day", y = "Frequency")

mean_steps <- mean(steps$steps, na.rm = TRUE)
median_steps <- median(steps$steps, na.rm = TRUE)

mean_steps
median_steps


# total_steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
# step_rec <- as.data.frame(rowsum(data$steps, data$date, reorder = TRUE, na.rm = TRUE))
# names(step_rec) <- c("Total_Steps")
# ggplot(step_rec, aes(x = Total_Steps)) + 
#   geom_histogram()
# mean(step_rec[,1])
# median(step_rec[,1])
```
## What is the average daily activity pattern?
```{r}
interval <- data %>%
  filter(!is.na(steps)) %>%
  group_by(interval) %>%
  summarize(steps = mean(steps))
ggplot(interval, aes(x=interval, y=steps)) +
  geom_line()

interval[which.max(interval$steps),]
```
The interval 835 has the highest count of steps on average with 206 steps.


## Imputing missing values
```{r}
sum(is.na(data$steps))
```
Missing values are 2304.


Replace NA by mean value
```{r}
data_full <- data
nas <- is.na(data_full$steps)
avg_interval <- tapply(data_full$steps, data_full$interval, mean, na.rm=TRUE, simplify=TRUE)
data_full$steps[nas] <- avg_interval[as.character(data_full$interval[nas])]
```

```{r}
sum(is.na(data_full$steps))
```
No more missing values now.


```{r}
steps_full <- data_full %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print

ggplot(steps_full, aes(x = steps)) +
  geom_histogram(binwidth = 1000) +
  labs(title = "Histogram of Steps per day, including missing values", x = "Steps per day", y = "Frequency")

mean_steps_full <- mean(steps_full$steps, na.rm = TRUE)
median_steps_full <- median(steps_full$steps, na.rm = TRUE)

mean_steps_full
median_steps_full
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
data_full <- mutate(data_full, weektype = ifelse(weekdays(data_full$date) == "Saturday" | weekdays(data_full$date) == "Sunday", "weekend", "weekday"))
data_full$weektype <- as.factor(data_full$weektype)
head(data_full)

interval_full <- data_full %>%
  group_by(interval, weektype) %>%
  summarise(steps = mean(steps))
ggplot(interval_full, aes(x=interval, y=steps, color = weektype)) +
  geom_line() +
  facet_wrap(~weektype, ncol = 1, nrow=2)
```

